---
title: "R visualization of epidemic in United States"
author: "Wanqi"
date: 2020-06-27T21:13:54-06:00
categories: ["R"]
tags: ["US", "COVID-19", "Epidemic"]
output:
  html_document:
    df_print: paged
mainfont: Palatino
fontsize: 13pt
subtitle: The characteristics of the COVID-19 outbreak in United States
always_allow_html: yes
---
```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Epidemic in United States

On the another side of the world, however, the disease has rapidly transmitted through the world, especailly United States. As compare to China, it does not implement lockdown and social distancing policy at the early stage of the epidemic. But it's true that, US's mass screening and medical system performance has shown the its amazing health-care resources, which is shown in my previous assignment. However, many deaths should be preventable, and there is even more arguable policy recently.

By May 21st, 2020, which is the time I write this assignment, the virus is still rapidly transimitted, with daily new diagnosis is ~ 20,000. However, all 50 states are now reopening. 

Many public health experts argued that, states should not reopen currently, because economic interest should always be less impartant than lifes. They think that government are making compromises to meet economic interest or to save lifes. 

However, on the other hand, I believe, reopening is also save lifes. Because it is a narrow-minded assumption that the epidemic kills people only by virus. The epidemic kill people in three ways: 

 - Deaths directly cost by the virus
 
 - Deaths of some chronic diseases, e.g., diabetes or cancer patients who has barriers in transportation or drug supply. 
 - Deaths of people attacked by financial problem, e.g., bussiness people or workers no longer have money to pay for their health and survival. 


We don't know the consequense of this reopening policy, maybe it kill more people, maybe it save lots of people. But at least we can have some insights from data. 


## Aims 
 - To visualize the trend of COVID-19 increase trend in US and US.
 - To further explore the data depends on the information in datasets


## Load packages
```{r}
require(tidyverse) 
require(leaflet)
require(leafletCN) 
if(require(rgeos)){
  require(rgeos)
} else{
  install.packages(("rgeos"))
}
require(rgeos)
require(maptools)
```


## Load data


```{r}
require("ncovr")
usd <- get_us()
us_states <- usd[['states_current']]
us_daily <- usd[['states_daily']]
#us_daily_l <- ncovr::get_us(ports = c("states_daily")) 
```


```{r}
us_daily$date2 <- as.Date(as.character(us_daily$date), format="%Y%m%d") 
```

## How's the mass screening now? How's the cumulative postive cases now? How's the deaths and recovered cases now?

```{r}
daily <-
  us_daily %>% group_by(date2) %>% summarise(
    `Positive from testing` = sum(positive, na.rm = TRUE),
    `Negative from testing` = sum(negative, na.rm = TRUE),
    Pending = sum(pending, na.rm = TRUE),
    `Current hospitalized patients` = sum(hospitalizedCurrently, na.rm = TRUE),
    `Current patients in ICU`= sum(inIcuCurrently, na.rm = TRUE),
     `Current patients on ventilator`= sum(onVentilatorCurrently, na.rm = TRUE),
    Recovered = sum(recovered, na.rm = TRUE),
    Death = sum(death, na.rm = TRUE)
  )
```

ok!
```{r}
us1<-daily %>%
  filter(date2>as.Date("2020-03-21")) %>% ggplot() + geom_point(aes(x = date2, y = `Positive from testing`, color = "Positive from testing")) + geom_point(aes(x = date2, y = `Negative from testing`, color = "Negative from testing")) + geom_point(aes(x = date2, y = Pending, color = "Pending")) + labs(y = "Cases", color = "Legend") + scale_colour_manual( values=c(`Positive from testing` = "red", `Negative from testing` = "green", Pending = "blue"))

```

```{r fig8, fig.cap="Results from 'mass screening' in U.S."}
plotly::ggplotly(us1)
```



```{r}
us2<-daily %>%
  filter(date2>as.Date("2020-03-21")) %>% ggplot() + geom_point(aes(x = date2, y = `Positive from testing`, color = "Diagnosed"),size=0.8) + geom_point(aes(x = date2, y = Death, color = "Death"),size=0.8) + geom_point(aes(x = date2, y = Recovered, color = "Recovered"),size=0.8) + labs(y = "Cases", color = "Legend") + scale_colour_manual( values=c(Diagnosed = "red", Death = "black", Recovered = "blue"))+
  geom_line(aes(x = date2, y = `Positive from testing`),size = 0.8, col="red")+ geom_line(aes(x = date2, y = Death),size = 0.8, col="black")+ geom_line(aes(x = date2, y = Recovered),size = 0.8, col="blue")

```

```{r fig9, fig.cap="Survival outcome of patients in U.S."}
plotly::ggplotly(us2)
```

Let's make a better one!
```{r}
diff(daily$date2)
```

```{r fig10, fig.cap="Survival outcome of patients in U.S."}
require(dygraphs) 
require(xts)
us3<-daily %>% select(`Positive from testing`, Death, Recovered) %>% xts(daily$date2) %>% dygraph() %>%
dyOptions(fillGraph = TRUE) %>% dyRangeSelector()
us3
```


## How's the status of patients who are in hospitals

So, currently, how's the patients are in hospital, how many of them are in ICU, how many of them are on ventilator?
```{r fig11, fig.cap="Status of patient in hospitals"}
us3<-daily %>% select(`Current hospitalized patients`, `Current patients in ICU`, `Current patients on ventilator`) %>% xts(daily$date2) %>% dygraph() %>%
dyOptions(fillGraph = TRUE) %>% dyRangeSelector()
us3
```

You can see that, interestingly, there is a gap for patients in ICU from April 26th to May 7th. 



## How's the daily increase?
```{r}
diff(daily$date2)
```

```{r}
#This is not a good way
#tnew<-daily[-nrow(daily),2] %>% pull(`Positive from testing`) 
#tnew1<-c(0,tnew)
#daily$PinYesterday<-tnew1
#daily$Pnew<-daily$`Positive from testing`-daily$PinYesterday
```

```{r}
daily$Pnew<-c(0,diff(daily$`Positive from testing`))
daily$Dnew<-c(0,diff(daily$Death))
```


```{r fig12, fig.cap="Daily increase number of COVID-19 in U.S."}
d1<-daily %>% ggplot() + geom_point(aes(x = date2, y = Pnew, color = "Daily increase"), size=0.5)+
  geom_point(aes(x = date2, y = Dnew, color = "Daily death"),size=0.5)+
  labs(y = "Cases", color = "Legend") + scale_colour_manual( values=c(`Daily increase` = "red", `Daily death` = "black"))+ geom_line(aes(x = date2, y = Dnew),size = 0.8)+ geom_line(aes(x = date2, y = Pnew),size = 0.8, col="red")

plotly::ggplotly(d1)
```

## How's the epidemic each state now?

```{r fig13, fig.cap="Total positive cases in U.S. currently"}
require(usmap)
mapus<-us_daily %>% filter(date==max(date)) %>%  plot_usmap("state", data = ., values = "positive") +
scale_fill_gradient(low = "green", high = "red", na.value = "white", trans = "log", limits = c(1, 10^6), breaks = 10 ^(0:6), labels = c('1', '10', '100', '1,000', '10,000', '100,000', '1000,000'))+ theme(legend.title = element_blank())
mapus
```



 


# Acknowledgement

Many thanks to ncovr package [@Zhao2020], by which make this assignmnet better than my previous assignment 5. And I get data in Jilin city and Jilin province from ncov2019 package [@Yu2020]. Time series also use packages of dygraphs [@Vanderkam2018], ggplot2 [@Wickham2020], plotly [@Sievert2020]. And the maps are made by following package: amap in leaflet [@Cheng2019], rgeos [@Bivand2020], and leafletCN [@Lang2020], beginr [@Zhao2019]

